---
title: "Final Project"
author: "Alison and Walter"
date: 2020-12-4
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE}
library(tidyverse)
```

# Details on the datasets  

<!-- -------------------------------------------------- -->

### Spotify dataset: 
To read the csv of the song feature data data, use this code:

```{r read-feature}
filename <- "./data/df_data.csv"
df_data <- read_csv(filename)
df_data

```


```{r - basic EDA}
df_data %>% select(week_position) %>% summary()
df_data %>% 

    arrange(desc(peak_position)) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = peak_position, y = week_position, color = weeks_on_chart)) +
  labs(
    x = "Maximum Position on Billboard Charts",
    y = "Where the song is on the chart"
    
  )
  
```




```{r}
df_data %>% 
  filter(Year == 2010 & Month ==1) %>%
  arrange(desc(week_position))%>%
  ggplot() + 
  geom_point(mapping = aes(x = danceability, y = week_position))

```

```{r - look at deancablility and peak_position in 2010}

df_data %>%
  filter(week_position == peak_position) %>%
  group_by(SongID) %>%
  slice(which.min(peak_position)) %>% #select the row from each song with the lowest peak_position.
  filter(Year == 2010) %>%
  ungroup() %>%
  drop_na(danceability) %>%
  mutate( dance_bin=cut_width(danceability, width=0.05, boundary=0) ) %>%
  ggplot(mapping = aes(x = dance_bin, y = peak_position)) +
  geom_boxplot() +
  coord_flip()

```
# Observations

the lowest danceability rate also has the worst rating for peak_position.
no linear relationship between danceability and peak_position except the bottom

```{r danceability and rank during all time}

df_data %>%
  filter(week_position == peak_position) %>%
  group_by(SongID) %>%
  slice(which.min(peak_position)) %>% #select the row from each song with the lowest peak_position.
  ungroup() %>%
  drop_na(danceability) %>%
  mutate( dance_bin=cut_width(danceability, width=0.075, boundary=0) ) %>%
  ggplot(mapping = aes(x = dance_bin, y = peak_position)) +
  geom_boxplot() +
  coord_flip()
```
 # Observations
 
 no real correlation
 
```{r}
df_data %>%
  filter(week_position == peak_position) %>%
  group_by(SongID) %>%
  slice(which.min(peak_position)) %>% #select the row from each song with the lowest peak_position.
  filter(Year == 2018) %>%
  ungroup() %>%
  drop_na(danceability) %>%
  mutate( dance_bin=cut_width(danceability, width=0.075, boundary=0) ) %>%
  ggplot(mapping = aes(x = dance_bin, y = peak_position)) +
  geom_boxplot() +
  coord_flip()
```
 # Observations
 only relationship for extremes
 
```{r - explore tempo and rank}

df_data %>%
  filter(week_position == peak_position) %>%
  group_by(SongID) %>%
  slice(which.min(peak_position)) %>% #select the row from each song with the lowest peak_position.
  ungroup() %>%
  drop_na(tempo) %>%
  mutate( tempo_bin=cut_width(tempo, width=15, boundary=0) ) %>%
  ggplot(mapping = aes(x = tempo_bin, y = peak_position)) +
  geom_boxplot() +
  coord_flip()

df_data %>%
  filter(week_position == peak_position) %>%
  group_by(SongID) %>%
  slice(which.min(peak_position)) %>% #select the row from each song with the lowest peak_position.
  ungroup() %>%
  drop_na(tempo) %>%
  filter(tempo >=226)
```
 look into these songs more
 
 
 
```{r}
df_data %>%
  filter(SongID == "HeartKenny Chandler") %>% 
  mutate(date = ISOdate(Year, Month, Day)) %>%
  
  ggplot(aes(x = date, y = week_position)) +
  geom_line()
  
df_data %>%
  filter(SongID == "TuffAce Cannon") %>% 
  mutate(date = ISOdate(Year, Month, Day)) %>%
  
  ggplot(aes(x = date, y = week_position)) +
  geom_line()
```
 
```{r explore}
df_data %>%
  filter(week_position == peak_position) %>%
  group_by(SongID) %>%
  slice(which.min(peak_position)) %>% #select the row from each song with the lowest peak_position.
  ungroup() %>%
  drop_na(time_signature) %>%
  mutate( bin=cut_width(time_signature, width=.1, boundary=0) ) %>%
  ggplot(mapping = aes(x = bin, y = peak_position)) +
  geom_boxplot() +
  coord_flip()
```

