---
title: "Final Report - Spotify and Billboard Rankings"
author: Alison and Walter Villa
date: 2020-12-15
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose: 

What song attributes lead towards success on the Billboard charts? 


## Background: 

Each year, new songs emerge and become some of the year’s top hits. Whether it's because of a viral video or a certain time of year, new songs seem to become popular overnight. The Billboard charts are a place in which songs and albums are ranked based on their weekly popularity in the United States. The charts can be ranked in categories such as sales, streams, or airplay. All three of these categories are used to compile the Hot 100 song chart, including data that come from YouTube and other video streaming sites.  

In 2005, Billboard made the change to include paid digital downloads from digital music retail stores like AmazonMP3 and iTunes. This meant that a song was able to make it onto the charts based solely on digital downloads. In 2007, it incorporated digital streams into the Top 100. Over the years, Billboard's methodology changed in ways that would allow different types of genres to gain a position on the charts through digital downloads and streaming plays. It's most recent change has had major impact in how songs were ranked on the billboard charts. This change had to do with YouTube video streaming data, which enhanced a formula that incorporates on-demand audio streaming and online radio streaming. Billboard made this change in order to reflect the diverse music consumption platforms of today's day and age. 

Spotify is a Swedish audio streaming and media services provider. Having come out in the early 2000's, it offers millions of digital copyright music and podcasts. While free service allows you to listen to songs alongside advertisements, paid subscriptions offer easier accessibility and more features for users. People from all around the world are able to listen to the latest trending songs on the platform. Interestingly enough, Spotify stores certain features for different songs, which is useful information for noticing certain trends in the type of songs that perform a certain way. On January 3, 2020, Billboard made the decision to count the popularity of official music videos on YouTube, as well as on streaming platforms like Apple Music, Spotify, Tidal, and Vevo into the Billboard 200, according to the New York Times. As a group, we wanted to investigate any potential patterns or trends in songs landing on the Billboard charts in correlation to Spotify's top tracks. 


## Data: 

Our data is composed of Spotify's top songs by year in the world and top songs ranked on the Billboard charts. The Spotify dataset contains several variables corresponding to each song's features (i.e genre, bpm [beats per minute], energy, liveness, etc) defined by Spotify itself. The Billboard dataset contains values from each track pulled from the Spotify Web API to be able to have information on each track listed on the Billboard charts.  



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
